Gametic Selection, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2017-01-24 LaTeX/Turnover.tex
sex symbols, notes in table 2

% Including pdf figures
\usepackage{graphicx}
\graphicspath{ {../Plots/} }
\usepackage{pdfpages}
%really place a figure in a location
\usepackage{float}
%Overrun caption
\usepackage[CaptionAfterwards]{fltpage}
% Math stuff
\usepackage{amsmath}
% Bibliographies
\usepackage[numbers]{natbib}
\bibpunct{(}{)}{,}{a}{}{;}

@@

\begin{abstract}
Sex determination systems are remarkably dynamic; many studied taxa display transitions of sex-determining genes between chromosomes or the evolution of new sex-determining systems. 
%Predominant theories in which new sex-determining systems are favoured by selection involve sex ratio selection or fitness differences between sexes (e.g., sexually antagonistic selection). 
Here, we utilize population genetic models to study the spread of novel sex-determiners in systems with 
%a period of selection during the haploid stage produced by one sex, e.g., pollen or sperm competition. 
haploid gametic selection, e.g., pollen or sperm competition.
Haploid selected loci experience a form of sex-specific selection (because gametic competition occurs predominantly among haploids produced by males) and can cause sex ratios at birth to become biased (because sex ratios are determined by the fertilization success of X- versus Y-bearing pollen/sperm). 
%We find that the evolution of sex determination systems where mothers determine sex at birth (e.g., environmental sex determination where sex is determined at birth) is influenced by classic Fisherian sex ratio selection. {\color{red}(Maybe not true???)}
%However, notably, we find that the spread of new genetic sex determination systems is not affected by sex ratio biases that are caused by gametic selection because sex ratios become biased after parental provisioning has occurred (even if pollen/sperm competition occurs within the mother). 
Notably, we find that the spread of new genetic sex determination systems is not affected by sex ratio biases that are caused by haploid competition or meiotic drive. 
In addition, we find that linkage of an ancestral sex chromosome to a locus under haploid selection can favour transitions between male and female heterogamety (e.g., XY to ZW), which is not the case for any forms of diploid sex specific selection (e.g., sexually antagonistic selection).
During these transitions, new sex-determining alleles spread despite breaking up favourable associations that build up between ancestral sex-determining loci and selected loci, reducing population mean fitness. 
Furthermore, a period of selection among haploids can favour the stable maintenance of polymorphic sex determination systems. 
Thus, our models offer several new insights to be explored as information about sex determination in non-model taxa accumulates.

@@

%life-cycle
In each generation, we census the genotype frequencies in male and female gametes/gametophytes (hereafter gametes) before haploid competition (see Sup.\ Mat. for recursion equations). 
First, competition occurs among male gametes (sperm/pollen competition) and among female gametes (egg/ovule competition). 
Selection during haploid competition depends on the \textbf{A} locus genotype, fitnesses are $w_{A}^m$ and $w_{a}^m$ for male gametes and $w_{A}^f$ and $w_{a}^f$ for female gametes, see table \ref{tab:fitnesstable}. 
Random mating then occurs between male and female gametes.
The resulting zygotes develop as males or females, depending on their genotypes at the \textbf{X} and \textbf{M} loci (and the \textbf{M} genotype of their mother in the case of maternal control) as described above.
Diploid males and females then experience selection, male fitness is given by $w_{h}^{m}$ and female fitness by $w_{h}^{f}$, where $h$ is the genotype at the \textbf{A} locus ($h \in AA, Aa, aa$).  
%Diploids compete with others of the same sex, where selection depends on the sex of the individual and its genotype at the \textbf{A} locus.
The next generation of gametes are then produced by meiosis, during which recombination and sex-specific meiotic drive can occur. 
Recombination occurs between loci \textbf{X} and \textbf{A} with probability $r$, between loci \textbf{A} and \textbf{M} with probability $R$, and between loci \textbf{X} and \textbf{M} with probability $\chi$.
Therefore, any order of the loci can be modelled with appropriate choices of $r$, $R$, and $\chi$ (see Table \ref{tab:chisubstitutions}). 
Males/females that are heterozygous at the \textbf{A} locus experience meiotic drive; $Aa$ heterozgotes of sex $d$ produce gametes bearing allele $A$ with probability $\alpha^d$. 
Thus, the \textbf{A} locus can experience sex-specific haploid competition, diploid selection and/or meiotic drive.

\begin{table}[ht]
\centering
\smallskip
\caption{Fitness of different genotypes in sex $d$}
\begin{tabular}{l c }
\hline\hline
  Genotype & Fitness during haploid competition \\ [0.5ex] \hline
  A & $w_{A}^d = 1+t^d$ \\
  a & $w_{a}^d = 1$ \\ [0.5ex] \hline
  Genotype & Fitness during diploid selection \\ [0.5ex] \hline
  AA & $w_{AA}^d = 1+ s^d$ \\
  Aa & $w_{Aa}^d = 1+h^k s^d$ \\
  aa & $w_{aa}^d = 1$ \\ [0.5ex] \hline
  Genotype & Tranmission during meiosis in $Aa$ heterozygotes \\ [0.5ex] \hline
  A & $\alpha^d=1/2+\alpha_{\Delta}^{d}/2$ \\
  a & $(1-\alpha^d)=1/2-\alpha_{\Delta}^{d}/2$ \\
  \hline \hline
  \label{tab:fitnesstable}
 \end{tabular}
\end{table}

@@

%I think this should be moved below because we have some 'results' where we don't assume weak selection but we only have the equilibrium calculated analytically for weak selection. 
\subsection*{Resident equilibrium and stability}
In the resident population (allele $M$ fixed), we follow the frequency of $A$ in female gametes (eggs) from an XX female, $p^f_X$, and in X-bearing, $p^m_X$, and Y-bearing, $p^m_Y$, male gametes (sperm). 
We also track the total frequency of Y-bearing male gametes (sperm), $q$, which may deviate from $1/2$ due to meiotic drive in males.

Various forms of selection can maintain a polymorphism at the \textbf{A} locus, including sexually antagonistic selection, overdominance and conflicts between diploid selection and selection upon haploid genotypes \citep[ploidally antagonistic selection,][]{Immler:2012tl} or a combination of these selective regimes. 
Here, we assume that selection and meiotic drive are weak relative to recombination ($s^k$, $t^k$, $\alpha_{\Delta}^k$ of order $\epsilon$). 
The maintenance of a polymorphism at the \textbf{A} locus then requires that

\begin{equation}
\begin{split}
0&<-((1-h^f)s^f +(1-h^m) s^m + t^f +t^m + \alpha_{\Delta}^f+\alpha_{\Delta}^m)\\
%
0&<(h^f s^f +h^m s^m + t^f +t^m + \alpha_{\Delta}^f+\alpha_{\Delta}^m).
\end{split}
\end{equation}

\noindent
which indicates that a polymorphism is maintained under various selective regimes. 
In particular special cases, e.g., no sex-differences in selection or meiotic drive ($s^m=s^f$, $h^m=h^f$, and $\alpha^m=\alpha^f=1/2$), the equilibrium allele frequency and stability can be calculated analytically without assuming weak selection. 
However, here, we focus on weak selection in order to make fewer assumptions about fitnesses.

Given that a polymorphism is maintained at the \textbf{A} locus by selection, with weak selection and drive, to leading order, the frequencies of $A$ in each type of gamete are the same ($\hat{p}^f_X=\hat{p}^m_X=\hat{p}^m_Y=\bar{p}$) and given by

\begin{equation}
\bar{p}=\frac{h^f s^f + h^m s^m +t^f+t^m+\alpha_{\Delta}^f+\alpha_{\Delta}^m}
{(2h^f-1)s^f+(2h^m-1)s^m}
+O(\epsilon)
.
\label{eq:pAve}
\end{equation}

@@

\begin{equation}
\begin{split}
\hat{p}^m_X-\hat{p}^f_X&=V_{A}\big{(}D^m - D^f +\alpha_{\Delta}^m-\alpha_{\Delta}^f \big{)}
+O(\epsilon^2)\\
%
\hat{p}^m_Y-\hat{p}^f_X&=V_{A}\big{(}D^m - D^f+\alpha_{\Delta}^m-\alpha_{\Delta}^f+(1-2r)(t^m-t^f)\big{)}/2r
+O(\epsilon^2)\\
%
\hat{p}^m_Y-\hat{p}^m_X&=V_{A}\big{(}D^m - D^f+\alpha_{\Delta}^m-\alpha_{\Delta}^f+t^m-t^f\big{)}(1-2r)/2r
+O(\epsilon^2)
\end{split}
\label{eq:freq_diffs}
\end{equation}

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $D^d=\big{(} \bar{p}s^d+(1-\bar{p})h^d s^d\big{)} -\big{(} \bar{p}h^d s^d+(1-\bar{p}) \big{)}$ corresponds to the difference in fitness between $A$ and $a$ alleles in diploids of sex $d$ ($\bar{p}$ is the leading-order probability of mating with an $A$-bearing gamete from the opposite sex). 
The frequency of Y among male gametes depends upon the difference in $A$ allele frequency on X- and Y-bearing male gametes and the strength of meiotic drive in favour of the $A$ allele in males, $q=1/2+\alpha_{\Delta}^m(\hat{p}^m_Y-\hat{p}^m_X)/2+O(\epsilon^3)$.
Without haploid competition or drive ($\alpha_{\Delta}^d=t^d=0$), these results reduce to those of \citet{vanDoorn:2007eu}.

@@

\noindent
where $b$ is the average of the growth rates of the two haplotypes that carry the $m$ allele ($mA$ and $ma$), $b=(\lambda_{mA}^{m}+\lambda_{ma}^{m})/2$, and $c$ also involves the fitness of $m$ alleles when they recombine onto the other \textbf{A} background in a heterozygote, $c=\lambda_{mA}\lambda_{ma}+\rho_{mA}\rho_{ma}$, see table \ref{tab:haplotype_growth}.

\begin{threeparttable}[ht]
\centering
\smallskip
\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W chromosomes}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=\{p_{X}^{f} w_{A}^{f} w_{A}^{m} w_{AA}^{m}+(1-p_{X}^{f}) w_{a}^{f} w_{A}^{m} w_{Aa}^{m}\alpha^m(1-R)\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^m$ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=\{(1-p_{X}^{f}) w_{a}^{f} w_{a}^{m} w_{aa}^{m}+p_{X}^{f} w_{A}^{f} w_{a}^{m} w_{Aa}^{m}(1-\alpha^m)(1-R)\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^m $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R \{(1-p_{X}^{f}) w_{a}^{f} w_{A}^{m} w_{Aa}^{m} (1-\alpha_{m})\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^m $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\{p_{X}^{f} w_{A}^{f} w_{a}^{m} w_{Aa}^{m} \alpha_{m}\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^m$\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  neo-W ($k=1$) \** \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=\{\bar{p}^{m} w_{A}^{m} w_{A}^{f} w_{AA}^{f}+(1-\bar{p}^{m}) w_{a}^{m} w_{A}^{f} w_{Aa}^{f}\alpha^f(1-R)\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^f$ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=\{(1-\bar{p}^{m}) w_{a}^{m} w_{a}^{f} w_{aa}^{f}+\bar{p}^{m} w_{A}^{m} w_{a}^{f} w_{Aa}^{f}(1-\alpha^f)(1-R)\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^f $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R \{(1-\bar{p}^{m}) w_{a}^{m} w_{A}^{f} w_{Aa}^{f} (1-\alpha_{f})\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^f $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\{\bar{p}^{m} w_{A}^{m} w_{a}^{f} w_{Aa}^{f} \alpha_{f}\}/ \bar{w}_{H}^{f} \bar{w}_{H}^{m} \bar{w}^f$\\ [1ex] 
  \hline \hline 
   \end{tabular}
      \begin{tablenotes}
      \scriptsize
      \item \** $\bar{p}^{m}=p_{Y}^{m}q+p_{X}^{m}(1-q)$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes (pollen/sperm). 
    \end{tablenotes}
  \label{tab:haplotype_growth}
\end{threeparttable}

Equation \eqref{eq:charpoly_neoY} and table \ref{tab:haplotype_growth} illustrate a number of key points about the invasion of neo-Y and neo-W mutations. 
For a neo-Y, invasion depends on the relative \textcolor{red}{(is this right, is relative fitness divided by mean fitness or difference from 1???)} fitness of $A$-bearing and $a$-bearing male gametes (i.e., in sperm only, not eggs). 
The fitness of male gametes partly depends on the allele carried by the female gamete that they mate with (e.g., $A$ with probability $p_{X}^{f} w_{A}^{f}/\bar{w}_{H}^{f}$).
Similarly, invasion of a neo-W depends on the relative fitness of $A$-bearing and $a$-bearing female gametes. 
However, in the case of a neo-W, the allele carried by the male gamete that they mate with can come from either an X-bearing or a Y-bearing sperm (e.g., $A$ with probability $\bar{p}^{m} w_{A}^{m}/\bar{w}_{H}^{m}$, where $\bar{p}^{m}=p_{Y}^{m}q+p_{X}^{m}(1-q)$). In either case, the zygote will then develop as a female due to the presence of a neo-W. 
By contrast, females that do not carry the neo-W, only result from matings with X-bearing sperm (e.g., matings with $A$-bearing sperm occur with probability $\bar{p}_{X}^{m} w_{A}^{m}/\bar{w}_{H}^{m}$). 
If the \textbf{A} locus is initially linked to the ancestral sex-determining locus, \textbf{X}, (i.e., $r<1/2$) the frequency of the $A$ allele among X- and Y-bearing sperm can differ (equation \ref{eq:freq_diffs}). 
Thus, eggs with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating with male gametes.

We are particularly concerned with whether or not a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, that solves \eqref{eq:charpoly_neoY} is greater than one. 
If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur (if $\left\{(\lambda_{mA}-1)+ (\lambda_{ma}-1) \right\}/2>0$, $\lambda>1$). 
If neither haplotype increases in frequency ($\lambda_{mA}, \lambda_{ma} < 1$ \textcolor{red}{is this notation valid?}), the $m$ allele will not invade. 
Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires
\begin{equation}\label{eq:neoYR}
R \left[ 
\frac{p^f_X w^m_a (1-\alpha^m)}{(\lambda_{mA}-1)\bar{w}_{H}^{m}} + 
\frac{(1 - p^f_X) w^m_A \alpha^m \bar{w}_{H}^{m}}{(\lambda_{ma}-1)\bar{w}_{H}^{m}} \right] 
\frac{w^m_{Aa}}{\bar{w}^m} < 1,
\end{equation}

@@

\begin{equation}
\lambda_{W',XY} =\lambda_{Y',XY}+\left(2\alpha_{\Delta}^m-2\alpha_{\Delta}^f+t^m-t^f \right) \left( \hat{p}^m_Y-\hat{p}^m_X \right)
+O\left(\epsilon^3 \right) \\
\label{eq:lambda_neoW}
\end{equation}

>>>>

% Including pdf figures
\usepackage{graphicx}
\graphicspath{ {../Plots/} }
\usepackage{pdfpages}
%really place a figure in a location
\usepackage{float}
%Overrun caption
\usepackage[CaptionAfterwards]{fltpage}
% Math stuff
\usepackage{amsmath}
\usepackage{stix}
% Bibliographies
\usepackage[numbers]{natbib}
\bibpunct{(}{)}{,}{a}{}{;}

@@

\begin{abstract}
Sex determination systems are remarkably dynamic; many studied taxa display transitions of sex-determining genes between chromosomes or the evolution of entirely new sex-determining systems. 
%Predominant theories in which new sex-determining systems are favoured by selection involve sex ratio selection or fitness differences between sexes (e.g., sexually antagonistic selection). 
Here, we utilize population genetic models to study the spread of novel sex-determiners in systems with 
%a period of selection during the haploid stage produced by one sex, e.g., pollen or sperm competition. 
haploid gametic selection, e.g., pollen or sperm competition.
Haploid selected loci experience a form of sex-specific selection (because gametic competition occurs predominantly among haploids produced by males) and can cause sex ratios at birth to become biased (because sex ratios are determined by the fertilization success of X- versus Y-bearing pollen/sperm). 
%We find that the evolution of sex determination systems where mothers determine sex at birth (e.g., environmental sex determination where sex is determined at birth) is influenced by classic Fisherian sex ratio selection. {\color{red}(Maybe not true???)}
%However, notably, we find that the spread of new genetic sex determination systems is not affected by sex ratio biases that are caused by gametic selection because sex ratios become biased after parental provisioning has occurred (even if pollen/sperm competition occurs within the mother). 
Notably, we find that the spread of new genetic sex determination systems is not affected by sex ratio biases that are caused by haploid competition or meiotic drive. 
In addition, we find that linkage of an ancestral sex chromosome to a locus under haploid selection can favour transitions between male and female heterogamety (e.g., XY to ZW), which is not the case for any forms of diploid sex specific selection (e.g., sexually antagonistic selection).
During these transitions, new sex-determining alleles spread despite breaking up favourable associations that build up between ancestral sex-determining loci and selected loci, reducing population mean fitness. 
Furthermore, a period of selection among haploids can favour the stable maintenance of polymorphic sex determination systems. 
Thus, our models offer several new insights to be explored as information about sex determination in non-model taxa accumulates.

@@

%life-cycle
In each generation, we census the genotype frequencies in male and female gametes/gametophytes (hereafter gametes) before haploid competition (see Sup.\ Mat. for recursion equations). 
First, competition occurs among male gametes (sperm/pollen competition) and among female gametes (egg/ovule competition) separately. 
Selection during haploid competition depends on the \textbf{A} locus genotype, relative fitnesses are given by $w_A^\Hermaphrodite$ and $w_a^\Hermaphrodite$ ($\Hermaphrodite \in \{\female,\male\}$; see table \ref{tab:fitnesstable}). %$w_{A}^\male$ and $w_{a}^\male$ for male gametes and $w_{A}^\female$ and $w_{a}^\female$ for female gametes, see table \ref{tab:fitnesstable}. 
Random mating then occurs between male and female gametes.
The resulting zygotes develop as males or females, depending on their genotypes at the \textbf{X} and \textbf{M} loci (and the \textbf{M} genotype of their mother in the case of maternal control) as described above.
Diploid males and females then experience selection, relative fitnesses are given by $w_{g}^{\male}$ in males and $w_{g}^{\female}$ in females, where $g$ is the diploid genotype at the \textbf{A} locus ($g \in \{AA, Aa, aa\}$).  
%Diploids compete with others of the same sex, where selection depends on the sex of the individual and its genotype at the \textbf{A} locus.
The next generation of gametes are then produced by meiosis, during which recombination and sex-specific meiotic drive can occur. 
Recombination occurs between loci \textbf{X} and \textbf{A} with probability $r$, between loci \textbf{A} and \textbf{M} with probability $R$, and between loci \textbf{X} and \textbf{M} with probability $\chi$.
Therefore, any order of the loci can be modelled with appropriate choices of $r$, $R$, and $\chi$ (see Table \ref{tab:chisubstitutions}). 
Males/females that are heterozygous at the \textbf{A} locus experience meiotic drive; $Aa$ heterozgotes of sex $\Hermaphrodite$ produce gametes bearing allele $A$ with probability $\alpha^\Hermaphrodite$. 
Thus, the \textbf{A} locus can experience sex-specific haploid competition, diploid selection and/or meiotic drive.

\begin{table}[ht]
\centering
\smallskip
\caption{Relative fitness of different genotypes in sex $\Hermaphrodite \in \{\female,\male\}$}
\begin{tabular}{l c }
\hline\hline
  Genotype & Relative fitness during haploid competition \\ [0.5ex] \hline
  A & $w_{A}^\Hermaphrodite = 1+t^\Hermaphrodite$ \\
  a & $w_{a}^\Hermaphrodite = 1$ \\ [0.5ex] \hline
  Genotype & Relative fitness during diploid selection \\ [0.5ex] \hline
  AA & $w_{AA}^\Hermaphrodite = 1+ s^\Hermaphrodite$ \\
  Aa & $w_{Aa}^\Hermaphrodite = 1+h^\Hermaphrodite s^\Hermaphrodite$ \\
  aa & $w_{aa}^\Hermaphrodite = 1$ \\ [0.5ex] \hline
  Genotype & Tranmission during meiosis in $Aa$ heterozygotes \\ [0.5ex] \hline
  A & $\alpha^\Hermaphrodite=1/2+\alpha_{\Delta}^{\Hermaphrodite}/2$ \\
  a & $(1-\alpha^\Hermaphrodite)=1/2-\alpha_{\Delta}^{\Hermaphrodite}/2$ \\
  \hline \hline
  \label{tab:fitnesstable}
 \end{tabular}
\end{table}

@@

%I think this should be moved below because we have some 'results' where we don't assume weak selection but we only have the equilibrium calculated analytically for weak selection. 
\subsection*{Resident equilibrium and stability}
In the resident population (allele $M$ fixed), we follow the frequency of $A$ in female gametes (eggs) from an XX female, $p^\female_X$, and in X-bearing, $p^\male_X$, and Y-bearing, $p^\male_Y$, male gametes (sperm).
We also track the total frequency of Y-bearing male gametes, $q$, which may deviate from $1/2$ due to meiotic drive in males.

Various forms of selection can maintain a polymorphism at the \textbf{A} locus, including sexually antagonistic selection, overdominance and conflicts between diploid selection and selection upon haploid genotypes \citep[ploidally antagonistic selection,][]{Immler:2012tl} or a combination of these selective regimes. 
Here, we assume that selection and meiotic drive are weak relative to recombination ($s^\Hermaphrodite$, $t^\Hermaphrodite$, $\alpha_{\Delta}^\Hermaphrodite$ of order $\epsilon$). 
The maintenance of a polymorphism at the \textbf{A} locus then requires that

\begin{equation}
\begin{split}
0&<-((1-h^\female)s^\female +(1-h^\male) s^\male + t^\female +t^\male + \alpha_{\Delta}^\female+\alpha_{\Delta}^\male)\\
%
0&<(h^\female s^\female +h^\male s^\male + t^\female +t^\male + \alpha_{\Delta}^\female+\alpha_{\Delta}^\male).
\end{split}
\end{equation}

\noindent
which indicates that a polymorphism is maintained under various selective regimes. 
In particular special cases, e.g., no sex-differences in selection or meiotic drive ($s^\male=s^\female$, $h^\male=h^\female$, and $\alpha^\male=\alpha^\female=1/2$), the equilibrium allele frequency and stability can be calculated analytically without assuming weak selection. 
However, here, we focus on weak selection in order to make fewer assumptions about fitnesses.

Given that a polymorphism is maintained at the \textbf{A} locus by selection, with weak selection and drive, to leading order, the frequencies of $A$ in each type of gamete are the same ($\hat{p}^\female_X=\hat{p}^\male_X=\hat{p}^\male_Y=\bar{p}$) and given by

\begin{equation}
\bar{p}=\frac{h^\female s^\female + h^\male s^\male +t^\female+t^\male+\alpha_{\Delta}^\female+\alpha_{\Delta}^\male}
{(2h^\female-1)s^\female+(2h^\male-1)s^\male}
+O(\epsilon)
.
\label{eq:pAve}
\end{equation}

@@

\begin{equation}
\begin{split}
\hat{p}^\male_X-\hat{p}^\female_X&=V_{A}\big{(}D^\male - D^\female +\alpha_{\Delta}^\male-\alpha_{\Delta}^\female \big{)}
+O(\epsilon^2)\\
%
\hat{p}^\male_Y-\hat{p}^\female_X&=V_{A}\big{(}D^\male - D^\female+\alpha_{\Delta}^\male-\alpha_{\Delta}^\female+(1-2r)(t^\male-t^\female)\big{)}/2r
+O(\epsilon^2)\\
%
\hat{p}^\male_Y-\hat{p}^\male_X&=V_{A}\big{(}D^\male - D^\female+\alpha_{\Delta}^\male-\alpha_{\Delta}^\female+t^\male-t^\female\big{)}(1-2r)/2r
+O(\epsilon^2)
\end{split}
\label{eq:freq_diffs}
\end{equation}

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $D^\Hermaphrodite=\big{(} \bar{p}s^\Hermaphrodite+(1-\bar{p})h^\Hermaphrodite s^\Hermaphrodite\big{)} -\big{(} \bar{p}h^\Hermaphrodite s^\Hermaphrodite+(1-\bar{p}) \big{)}$ corresponds to the difference in fitness between $A$ and $a$ alleles in diploids of sex $\Hermaphrodite \in \{\female,\male\}$ ($\bar{p}$ is the leading-order probability of mating with an $A$-bearing gamete from the opposite sex). 
The frequency of Y among male gametes depends upon the difference in $A$ allele frequency on X- and Y-bearing male gametes and the strength of meiotic drive in favour of the $A$ allele in males, $q=1/2+\alpha_{\Delta}^\male(\hat{p}^\male_Y-\hat{p}^\male_X)/2+O(\epsilon^3)$.
Without haploid competition or drive ($\alpha_{\Delta}^\Hermaphrodite=t^\Hermaphrodite=0$), these results reduce to those of \citet{vanDoorn:2007eu}.

@@

\noindent
where $b$ is the average of the growth rates of the two haplotypes that carry the $m$ allele ($mA$ and $ma$), $b=(\lambda_{mA}+\lambda_{ma})/2$, and $c$ also involves the fitness of $m$ alleles when they recombine onto the other \textbf{A} background in a heterozygote, $c=\lambda_{mA}\lambda_{ma}+\rho_{mA}\rho_{ma}$ (see table \ref{tab:haplotype_growth}).

\begin{threeparttable}[ht]
\centering
\smallskip
\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W chromosomes}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=\{p_{X}^{\female} w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male}+(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male}\alpha^\male(1-R)\}/ \{ \bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\male$ \} \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=\{(1-p_{X}^{\female}) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male}+p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1-\alpha^\male)(1-R)\}/ \{ \bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\male \}$ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R \{(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} (1-\alpha_{m})\}/ \{ \bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\male \} $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\{p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} \alpha_{m}\}/ \{ \bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\male \}$\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA}=\{\bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}\alpha^\female(1-R)\}/ \{ \bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\female \}$ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma}=\{(1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha^\female)(1-R)\}/ \{\bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\female \}$ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA}=R \{(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} (1-\alpha_{f})\}/ \{\bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\female \} $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma}=R\{\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} \alpha_{f}\}/ \{\bar{w}_{H}^{\female} \bar{w}_{H}^{\male} \bar{w}^\female \}$\\ [1ex] 
  \hline \hline 
   \end{tabular}
      \begin{tablenotes}
      \scriptsize
      \item $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes 
      \item $\bar{w}_H^\female = p_X^\female w_A^\female + (1-p_X^\female) w_a^\female$ is the mean fitness of female gametes
      \item $\bar{w}_H^\male = \bar{p}^{\male} w_A^\male + (1-\bar{p}^{\male}) w_a^\male$ is the mean fitness of male gametes
       \item $\bar{w}^\female = \{ p_X^\female w_A^\female (1-q) p_X^\male w_A^\male w_{AA}^\female + (1 - p_X^\female) w_a^\female (1-q) p_X^\male w_A^\male w_{Aa}^\female + p_X^\female w_A^\female (1-q) (1 - p_X^\male) w_a^\male w_{Aa}^\female + (1-p_X^\female) w_a^\female (1-q) (1 - p_X^\male) w_a^\male w_{aa}^\female \} / \{ \bar{w}_H^\female \bar{w}_H^\male\}$ is the mean fitness of females
       \item $\bar{w}^\male = \{ p_X^\female w_A^\female q p_Y^\male w_A^\male w_{AA}^\male + (1 - p_X^\female) w_a^\female q p_Y^\male w_A^\male w_{Aa}^\male + p_X^\female w_A^\female q (1 - p_Y^\male) w_a^\male w_{Aa}^\male + (1-p_X^\female) w_a^\female q (1 - p_Y^\male) w_a^\male w_{aa}^\male \} / \{ \bar{w}_H^\male \bar{w}_H^\male\}$ is the mean fitness of males
    \end{tablenotes}
  \label{tab:haplotype_growth}
\end{threeparttable}

Equation \eqref{eq:charpoly_neoY} and table \ref{tab:haplotype_growth} illustrate a number of key points about the invasion of neo-Y and neo-W mutations. 
For a neo-Y, invasion depends on the relative \textcolor{red}{(is this right, is relative fitness divided by mean fitness or difference from 1???)} fitness of $A$-bearing and $a$-bearing male gametes (i.e., in sperm only, not eggs). 
The fitness of male gametes partly depends on the allele carried by the female gamete that they mate with (e.g., $A$ with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$).
Similarly, invasion of a neo-W depends on the relative fitness of $A$-bearing and $a$-bearing female gametes. 
However, in the case of a neo-W, the allele carried by the male gamete that they mate with can come from either an X-bearing or a Y-bearing sperm (e.g., $A$ with probability $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$). In either case, the zygote will then develop as a female due to the presence of a neo-W. 
By contrast, females that do not carry the neo-W, only result from matings with X-bearing sperm (e.g., matings with $A$-bearing sperm occur with probability $\bar{p}_{X}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$). 
If the \textbf{A} locus is initially linked to the ancestral sex-determining locus, \textbf{X}, (i.e., $r<1/2$) the frequency of the $A$ allele among X- and Y-bearing sperm can differ (equation \ref{eq:freq_diffs}). 
Thus, eggs with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating with male gametes.

We are particularly concerned with whether or not a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, that solves \eqref{eq:charpoly_neoY} is greater than one. 
If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur (if $\left\{(\lambda_{mA}-1)+ (\lambda_{ma}-1) \right\}/2>0$, $\lambda>1$). 
If neither haplotype increases in frequency ($\lambda_{mA}, \lambda_{ma} < 1$ \textcolor{red}{is this notation valid?}), the $m$ allele will not invade. 
Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires
\begin{equation}\label{eq:neoYR}
R \left[ 
\frac{p^\female_X w^\male_a (1-\alpha^\male)}{(\lambda_{mA}-1)\bar{w}_{H}^{\male}} + 
\frac{(1 - p^\female_X) w^\male_A \alpha^\male \bar{w}_{H}^{\male}}{(\lambda_{ma}-1)\bar{w}_{H}^{\male}} \right] 
\frac{w^\male_{Aa}}{\bar{w}^\male} < 1,
\end{equation}

@@

\begin{equation}
\lambda_{W',XY} =\lambda_{Y',XY}+\left(2\alpha_{\Delta}^\male-2\alpha_{\Delta}^\female+t^\male-t^\female \right) \left( \hat{p}^\male_Y-\hat{p}^\male_X \right)
+O\left(\epsilon^3 \right) \\
\label{eq:lambda_neoW}
\end{equation}
